Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation

### 2014-10-17 writedown/osci.ele/submit.tex
rename

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field~\cite{foloppe2000all}, and dissolved in 644 TIP3P~\cite{jorgensen1983comparison}
water molecules. The grid-based energy correction map (CMAP)~\cite{mackerell2004extending}
for the backbone dihedral angles is also used.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS~4.5~\cite{pronk2013gromacs} together with the CHARMM
force field~\cite{bjelkmar2010implementation}.
First, an equilibrium NVT simulation at 300~K of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme (standard Gromacs implementation) with the aforementioned nonequilibrium
temperature control technique.  The
time step was $0.002~\textsf{ps}$.
The short-range
interaction (van der Waals interaction)
have a cut-off radius of 1.0~nm, and has been smoothed from $0.8$ to $1.0$~nm by the
``switch'' method provided by the GROMACS code.
A energy conserving Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm}
method was applied to calculate the electrostatic interaction in this
periodic system. For the direct space part of PME the cut-off and smoothing follow the same principles as those applied to the van der Waals interaction.
All hydrogen involving bonds are constrained by LINCS~\cite{hess1997lincs}, except 
the TIP3P water molecules which are constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
In all testing cases, this local thermostat is able to control
the system at the desired temperature, i.e.~300~K (the results
are not presented in this paper).
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic}~(in standard Gromacs implementation) with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
% {The use of leap-frog scheme and Parrinello-Rahman barostat
% was not our choice, but the default one of Gromacs.}
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

\begin{figure}
  \centering
  \includegraphics[width=1.00\textwidth]{fig-all.eps}
  % \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{c-2.eps}}
  % \includegraphics[width=0.37\textwidth]{fig-begin-1.eps}
  % \includegraphics[width=0.37\textwidth]{fig-end-1.eps}
  \caption{
    The probability density of the conformations of the alanine dipeptide plotted on the Ramachandran plot.
    (a): Two dihedral angles
    $\phi$ and $\psi$ are used to represent the molecular conformations of the alanine dipeptide.
    (b): The equilibrium probability density 
    at $t=0$~\textsf{ps}. (c): The probability density of equilibrium at $t=1000$~\textsf{ps}, resulting from the application of a constant EF of 1~V/nm. Darker color
    indicates higher probability.
    Plots are divided into 5 subregions corresponding
    to 5 molecular conformations:
    $\confaa = \{(\phi, \psi) | -134^\circ \leq \phi <  13^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $\confab = \{(\phi, \psi) |  128^\circ \leq \phi < 226^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $\confba = \{(\phi, \psi) | -110^\circ \leq \phi <  13^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $\confbb = \{(\phi, \psi) |  128^\circ \leq \phi < 250^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $\confc   = \{(\phi, \psi) |   13^\circ \leq \phi < 128^\circ, -180^\circ \leq \psi < 180^\circ\}$.
    Considering the angular periodicity, the angle, for example, $226^\circ$ corresponds to $-134^\circ$
    on the plot (b) and (c).
    % Considering the angular periodicity, if we write, for example,
    % $128^\circ \leq \phi <-134^\circ$, it implies that $\phi$ is in a range of $98$ degrees, between $128^\circ$ and $226^\circ$ ($226^\circ$ correspond to $-134^\circ$).
    This division is the same for both plot (b) and (c).
  }
  \label{fig:tmp4}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-1.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-2.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-3.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-4.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-5.eps}
  \caption{
    The net probability flux for the constant EF case.
    The warm-up time is $t_{warm} = 10$~ps.
    The probability flux $F_{J,I}(t)$ (defined by Eqn.~\eqref{eqn:tmp7}) is reported in units of $\textrm{ps}^{-1}$.
    From top, left to right we report $I = \confaa$, $\confab$, $\confba$; bottom, left to right $\confbb$ and
    $\confc$, respectively. In each plot, the red line stands for $J=\confaa$,
    green for $J=\confab$, dark blue for $J=\confba$, purple for $J=\confbb$ and light blue
    for $J=\confc$.
    {This plot is drawn from the four-branching-trajectories case (see the text).}
    }
  \label{fig:tmp6}
\end{figure}
The time-dependent probability density of the system to stay in a certain
conformation is given in
Fig.~\ref{fig:tmp5}.
In order to show that  our simulation results are robust with respect to the randomness introduced by the (local) Langevin thermostat, we have tested the effects of changing random seeds for the thermostat in the thermostated region. We compared simulation results obtained by starting a branching trajectory
from each initial conformation with those obtained from four different branching trajectories from each initial conformation (each of four with
different random seed for the Langevin thermostat in the thermostated region).
Results are consistent within the statistical uncertainty (see Fig.\ref{fig:tmp5}).
The $\beta$-sheet conformations $\confba$ and $\confbb$,
which are present at the initial stage after EF is switched on,
{fade away as the system relaxes}. The
probability of the $\alpha$-helix conformation $\confaa$, grows from 0.45 to 0.69.
The $\alpha$-helix $\confab$  does not change significantly  under the action of EF.
The left-hand helix conformation $\confc$ noticeably grows from 0 to 0.22.
We plot the net probability fluxes {of the four-branching-trajectories case} in Fig.~\ref{fig:tmp6}.
Results show that for short $t$, the net fluxes are generally non-zero,
however, 
as $t$ goes to infinity,
all fluxes converge to zero. This indicates
that as the EF is switched on, the system is driven away from the initial
equilibrium (at zero EF),  starting a dynamical nonequilibrium process.
After sufficiently long time, the system
is fully relaxed to the new stationary state, in
which it remains as long as the EF is switched on.
Of course we cannot exclude very slow nonequilibrium processes,
which cannot be captured by the duration of
our nonequilibrium simulation (1~ns), since it is
obviously that the proposed nonequilibrium approach
can only study time-dependent behaviours that are shorter than
the total time of the branching trajectories. 
% After a time long enough, the system
% is fully relaxed to the new equilibrium defined by the action of the constant EF,
% and stays there as long as the EF is kept.
An important point is that despite the system is finally relaxed to the new conformation,
the relaxation timescales corresponding to each conformation are rather different. 
From Fig.~\ref{fig:tmp5} and \ref{fig:tmp6}, we observe mainly
three different timescales 10~ps, 100~ps and 500~ps,
whose corresponding conformations are summarized in Tab.~\ref{tab:tmp1}.
This is a relevant result because it shows the possibility of employing an external EF as a tool to identify relevant time scales in the conformational behaviour of a molecule in solution.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{fig-rdf.eps}
  \includegraphics[width=0.48\textwidth]{fig-hbond.eps}
  \caption{The solvent structure around the alanine dipeptide.  
    (a): the radial distribution function (RDF) between the
    $\alpha$-carbon and the center-of-mass (COM) of the water molecules at
    $t = 0$, 10, 100 and 1000~ps.  (b): the number of hydrogen
    bonds formed between the alanine and water, and between two water
    molecules.  The number of hydrogen bonds between two water
    molecules is normalized by the number of water molecules in the
    system. 
  }
  \label{fig:tmp7a}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% solvent effect
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
So far we have focused our attention on the behaviour of the alanine
under the action of the EF. However, it is also important to understand
the effects of the EF on the solvating water molecules. Being water a
polar molecule the electrostatic interaction with an EF can
dramatically change its solvation structure around the alanine and
thus, in turn, influence its conformational behaviour. For this
reason, in this section we analyze the behaviour of the solvent around
the alanine under the action of the EF by investigating
the radial distribution function (RDF) between $\alpha$-carbon and the
center-of-mass (COM) of the water molecules
(see Fig.~\ref{fig:tmp7a} (a)). The average
RDF does not change significantly as a function of time
(within the statistical error). 
Fig.~\ref{fig:tmp7a}~(b) presents the number of hydrogen bonds between the
alanine and water molecules, and between two neighboring water molecules.
The number of hydrogen bonds are calculated by standard GROMACS
routine {\texttt {g\_hbond}}, with donor-acceptor distance cut-off 0.3~nm,
and hydrogen-donor-acceptor angle cut-off $20^{\circ}$.
We observe that the
number of hydrogen bonds remains almost the same when the EF is turned on,
although molecular dipoles tend to align along the direction of the EF (result
not shown). 
The EF-induced solvent effect on the alanine can be also accounted by calculating the
difference of the water-alanine electrostatic interaction 
without and with the EF.
The energy difference is found to be $-19.1$~kJ/mol that is
comparable to the electrostatic energy associated to the dipole of the alanine itself.
Therefore, we conclude that the solvation structure around the
alanine does not change or contribute significantly in determining the conformational dynamics of the alanine under the action of EF.

@@

\section{Case II:
  periodically oscillatory EF}
A step forward in our study is to consider a periodically oscillatory EF which has
a $\sin$-wave shape:
\begin{align}
  \vect E(t) = (E_0\sin(2\pi t / \period), 0, 0),
\end{align}
where $E_0$ is the intensity of the field, which is chosen to be
1.0~V/nm.  $\period$ is the oscillating period.
Here we tested three different
periods 10, 40 and 200~ps.
\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{fig-mode2-dipol.eps}
  \includegraphics[width=0.48\textwidth]{fig-mode2-order.eps}
  \caption{Illustration of the effect of the oscillatory EF on the system:
    (a) the dipole moment  and (b) the
    orientational order parameter of alanine dipeptide as a
    function of time. In (a),
    only the $x$-component of the dipole moment is
    shown. The black line refers to the static EF which is turned on
    at time $t_{warm} = 10$~ps. The red line refers to the oscillatory field with period of
    10~ps. The green line is of 40~ps; and the blue of 200~ps.}
  \label{fig:tmp8}
\end{figure}
Fig.~\ref{fig:tmp8} (a)
presents the $x$-component of the molecular dipole moment as
a function of time. The red, green and blue lines report the
results of $\period =10$~ps, 40~ps and 200~ps, respectively. The black line
shows the dipole moment under constant EF
for reference.
It shows that the oscillating period of the molecular dipole moment
is the same as the period of the oscillatory EF. Therefore,
the molecular dipole moment
is able to respond to the external EF almost immediately.
At $\period =200$~ps, the maximum magnitude of the molecular
dipole moment is close to the value obtained in the case of constant EF, which
means that the variation of the oscillatory EF is so slow that the
alanine has enough time to 
relax its dipole. However,
for $\period =10$ and 40~ps, the EF oscillates faster and the molecule 
does not have enough time to fully relax the dipole, as a consequence the maximum dipole moment is smaller than the previous case.
Fig.~\ref{fig:tmp8}~(b) presents the orientational order
parameter. The notations are the same as Fig~\ref{fig:tmp8}~(a).
For all cases the order parameter is much
smaller than in the case of constant EF. One possible reason is that the
relaxation of the order parameter is very slow, and the molecule in the case of oscillatory field is
not exposed to a strong enough EF for a time long enough.
For  $\period =10$ and 40~ps, the orientation of the alanine shows a {very} weak tendency
to be parallel to the EF. It is worth to notice that for constant EF,
we also observe a quick alignment of the molecular orientation vector to the EF
on the time scale of 10~ps. For $\period =200$~ps, the molecule tends periodically
to be more perpendicular to the EF, but this directional preference is much
weaker than in the case of constant EF.

\begin{figure}
  \centering
  \includegraphics[width=0.32\textwidth]{fig-meta-mode2-0010.eps}
  \includegraphics[width=0.32\textwidth]{fig-meta-mode2-0040.eps}
  \includegraphics[width=0.32\textwidth]{fig-meta-mode2-0200.eps}
  % \includegraphics[width=0.49\textwidth]{fig-meta-mode2-2000.eps}
  \caption{The probability of being in a given of conformation for alanine in a   periodically oscillatory EF. Different periods,
    i.e. $\period =10$~ps, 40~ps and 200~ps, are considered here.  The
    probability of being in a certain conformation $I$, which is
    denoted by $P_I$ in this paper, is plotted by colored lines against
    time. Red line: $I = \confaa$,  green line: $I = \confab$,  dark
    blue line: $I = \confba$,  pink line: $I = \confbb$ and  light blue
    line: $I = \confc$. For $\period =10$~ps and 40~ps, the nonequilibrium
    simulations last 1000~ps, while for $\period =200$~ps,
    they last 6400~ps.  }
  \label{fig:tmp9}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-1.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-2.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-3.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-4.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-5.eps}\\
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-1.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-2.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-3.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-4.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-5.eps}\\
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-1.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-2.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-3.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-4.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-5.eps}\\
  \caption{
    Integrated probability flux $Q_{J,I}(t)$ for the periodically
    oscillatory EF, is plotted against time (horizontal axis, in picoseconds).
    The integrated
    probability flux is defined by $Q_{J,I}(t) = \int_0^t F_{J,I}(\tau) d\tau$, where
    $F_{J,I}(t)$ is the net probability flux from conformation $J$ to $I$.
    See also Eqn.~\eqref{eqn:tmp7} and~\eqref{eqn:tmp7a} for definitions
    of $F_{J,I}(t)$ and $Q_{J,I}(t)$, respectively.
    From top to bottom, rows report results with  period 10, 40 and 200~ps, respectively.
    From left to right, the five
    columns show the integrated flux $Q_{J,\confaa}$, $Q_{J,\confab}$,
    $Q_{J,\confba}$, $Q_{J,\confbb}$ and $Q_{J,\confc}$, respectively. In each plot,
    the red line stands for $J=\confaa$, green for $J=\confab$, dark blue for $J=\confba$,
    purple for $J=\confbb$ and light blue for $J=\confc$.
    An increasing value of $Q_{J,I}$ indicates a net probability
      flux from conformation $J$ to $I$,
      while a decreasing value of $Q_{J,I}$ indicates a net probability
      flux from $I$ to $J$.
  }
  \label{fig:tmp10}
\end{figure}

>>>>

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field~\cite{foloppe2000all}, and dissolved in 644 TIP3P~\cite{jorgensen1983comparison}
water molecules. The grid-based energy correction map (CMAP)~\cite{mackerell2004extending}
for the backbone dihedral angles is also used.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS~4.5~\cite{pronk2013gromacs} together with the CHARMM
force field~\cite{bjelkmar2010implementation}.
First, an equilibrium NVT simulation at 300~K of
1000~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
0.5~\textsf{ns} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme (standard Gromacs implementation) with the aforementioned nonequilibrium
temperature control technique.  The
time step was $0.002~\textsf{ps}$.
The short-range
interaction (van der Waals interaction)
had a cut-off radius of 1.0~nm, and had been smoothed from $0.8$ to $1.0$~nm by the
``switch'' method provided by the GROMACS code.
A energy conserving Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm}
method was applied to calculate the electrostatic interaction in this
periodic system. For the direct space part of PME the cut-off and smoothing follow the same principles as those applied to the van der Waals interaction.
All hydrogen involving bonds were constrained by LINCS~\cite{hess1997lincs}, except 
the TIP3P water molecules which were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
In all testing cases, this local thermostat is able to control
the system at the desired temperature, i.e.~300~K (the results
are not presented in this paper).
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic}~(in standard Gromacs implementation) with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
% {The use of leap-frog scheme and Parrinello-Rahman barostat
% was not our choice, but the default one of Gromacs.}
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

\begin{figure*}
  \centering
  \includegraphics[width=1.00\textwidth]{fig-all.eps}
  % \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{c-2.eps}}
  % \includegraphics[width=0.37\textwidth]{fig-begin-1.eps}
  % \includegraphics[width=0.37\textwidth]{fig-end-1.eps}
  \caption{
    The probability density of the conformations of the alanine dipeptide plotted on the Ramachandran plot.
    (a): Two dihedral angles
    $\phi$ and $\psi$ are used to represent the molecular conformations of the alanine dipeptide.
    (b): The equilibrium probability density 
    at $t=0$~\textsf{ps}. (c): The probability density of equilibrium at $t=1000$~\textsf{ps}, resulting from the application of a constant EF of 1~V/nm. Darker color
    indicates higher probability.
    Plots are divided into 5 subregions corresponding
    to 5 molecular conformations:
    $\confaa = \{(\phi, \psi) | -134^\circ \leq \phi <  13^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $\confab = \{(\phi, \psi) |  128^\circ \leq \phi < 226^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $\confba = \{(\phi, \psi) | -110^\circ \leq \phi <  13^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $\confbb = \{(\phi, \psi) |  128^\circ \leq \phi < 250^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $\confc   = \{(\phi, \psi) |   13^\circ \leq \phi < 128^\circ, -180^\circ \leq \psi < 180^\circ\}$.
    Considering the angular periodicity, the angle, for example, $226^\circ$ corresponds to $-134^\circ$
    on the plot (b) and (c).
    % Considering the angular periodicity, if we write, for example,
    % $128^\circ \leq \phi <-134^\circ$, it implies that $\phi$ is in a range of $98$ degrees, between $128^\circ$ and $226^\circ$ ($226^\circ$ correspond to $-134^\circ$).
    This division is the same for both plot (b) and (c).
  }
  \label{fig:tmp4}
\end{figure*}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.22\textwidth]{fig-trans-010-flux-1.eps}
  \includegraphics[width=0.22\textwidth]{fig-trans-010-flux-2.eps}\\
  \vskip .3cm
  \includegraphics[width=0.22\textwidth]{fig-trans-010-flux-3.eps}
  \includegraphics[width=0.22\textwidth]{fig-trans-010-flux-4.eps}\\
  \vskip .3cm
  \includegraphics[width=0.22\textwidth]{fig-trans-010-flux-5.eps}
  \caption{
    The net probability flux for the constant EF case.
    The warm-up time is $t_{warm} = 10$~ps.
    The probability flux $F_{J,I}(t)$ (defined by Eqn.~\eqref{eqn:tmp7}) is reported in units of $\textrm{ps}^{-1}$.
    From top to bottom, left to right we report $I = \confaa$, $\confab$, $\confba$, $\confbb$ and
    $\confc$, respectively. In each plot, the red line stands for $J=\confaa$,
    green for $J=\confab$, dark blue for $J=\confba$, purple for $J=\confbb$ and light blue
    for $J=\confc$.
    {This plot is drawn from the four-branching-trajectories case (see the text).}
    }
  \label{fig:tmp6}
\end{figure}
The time-dependent probability density of the system to stay in a certain
conformation is given in
Fig.~\ref{fig:tmp5}.
In order to show that  our simulation results are robust with respect to the randomness introduced by the (local) Langevin thermostat, we have tested the effects of changing random seeds for the thermostat in the thermostated region. We compared simulation results obtained by starting a branching trajectory
from each initial conformation with those obtained from four different branching trajectories from each initial conformation (each of four with
different random seed for the Langevin thermostat in the thermostated region).
Results are consistent within the statistical uncertainty (see Fig.\ref{fig:tmp5}).
The $\beta$-sheet conformations $\confba$ and $\confbb$,
which are present at the initial stage after EF is switched on,
{fade away as the system relaxes}. The
probability of the $\alpha$-helix conformation $\confaa$, grows from 0.43 to 0.65.
The $\alpha$-helix $\confab$  does not change significantly  under the action of EF.
The left-hand helix conformation $\confc$ noticeably grows from 0.05 to 0.25.
We plot the net probability fluxes {of the four-branching-trajectories case} in Fig.~\ref{fig:tmp6}.
Results show that for short $t$, the net fluxes are generally non-zero,
however, 
as $t$ goes to infinity,
all fluxes converge to zero. This indicates
that as the EF is switched on, the system is driven away from the initial
equilibrium (at zero EF),  starting a dynamical nonequilibrium process.
After sufficiently long time, the system
is fully relaxed to the new stationary state, in
which it remains as long as the EF is switched on.
Of course we cannot exclude very slow nonequilibrium processes,
which cannot be captured by the duration of
our nonequilibrium simulation (1~ns), since it is
obviously that the proposed nonequilibrium approach
can only study time-dependent behaviours that are shorter than
the total time of the branching trajectories. 
% After a time long enough, the system
% is fully relaxed to the new equilibrium defined by the action of the constant EF,
% and stays there as long as the EF is kept.
An important point is that despite the system is finally relaxed to the new conformation,
the relaxation timescales corresponding to each conformation are rather different. 
From Fig.~\ref{fig:tmp5} and \ref{fig:tmp6}, we observe mainly
three different timescales 10~ps, 100~ps and 500~ps,
whose corresponding conformations are summarized in Tab.~\ref{tab:tmp1}.
This is a relevant result because it shows the possibility of employing an external EF as a tool to identify relevant time scales in the conformational behaviour of a molecule in solution.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig-rdf.eps}
  \includegraphics[width=0.4\textwidth]{fig-hbond.eps}
  \caption{The solvent structure around the alanine dipeptide.  
    (a): the radial distribution function (RDF) between the
    $\alpha$-carbon and the center-of-mass (COM) of the water molecules at
    $t = 0$, 10, 100 and 1000~ps.  (b): the number of hydrogen
    bonds formed between the alanine and water, and between two water
    molecules.  The number of hydrogen bonds between two water
    molecules is normalized by the number of water molecules in the
    system. 
  }
  \label{fig:tmp7a}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% solvent effect
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
So far we have focused our attention on the behaviour of the alanine
under the action of the EF. However, it is also important to understand
the effects of the EF on the solvating water molecules. Being water a
polar molecule the electrostatic interaction with an EF can
dramatically change its solvation structure around the alanine and
thus, in turn, influence its conformational behaviour. For this
reason, in this section we analyze the behaviour of the solvent around
the alanine under the action of the EF by investigating
the radial distribution function (RDF) between $\alpha$-carbon and the
center-of-mass (COM) of the water molecules
(see Fig.~\ref{fig:tmp7a} (a)). The average
RDF does not change significantly as a function of time
(within the statistical error). 
Fig.~\ref{fig:tmp7a}~(b) presents the number of hydrogen bonds between the
alanine and water molecules, and between two neighboring water molecules.
The number of hydrogen bonds are calculated by standard GROMACS
routine {\texttt {g\_hbond}}, with donor-acceptor distance cut-off 0.3~nm,
and hydrogen-donor-acceptor angle cut-off $20^{\circ}$.
We observe that the
number of hydrogen bonds remains almost the same when the EF is turned on,
although molecular dipoles tend to align along the direction of the EF (result
not shown). 
The EF-induced solvent effect on the alanine can be also accounted by calculating the
difference of the water-alanine electrostatic interaction 
without and with the EF.
The energy difference is found to be $-19.1$~kJ/mol that is
comparable to the electrostatic energy associated to the dipole of the alanine itself.
Therefore, we conclude that the solvation structure around the
alanine does not change or contribute significantly in determining the conformational dynamics of the alanine under the action of EF.

@@

\section{Case II:
  periodically oscillatory EF}
A step forward in our study is to consider a periodically oscillatory EF which has
a $\sin$-wave shape:
\begin{align}
  \vect E(t) = (E_0\sin(2\pi t / \period), 0, 0),
\end{align}
where $E_0$ is the intensity of the field, which is chosen to be
1.0~V/nm.  $\period$ is the oscillating period.
Here we tested three different
periods 10, 40 and 200~ps.
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig-mode2-dipol.eps}
  \includegraphics[width=0.4\textwidth]{fig-mode2-order.eps}
  \caption{Illustration of the effect of the oscillatory EF on the system:
    (a) the dipole moment  and (b) the
    orientational order parameter of alanine dipeptide as a
    function of time. In (a),
    only the $x$-component of the dipole moment is
    shown. The black line refers to the static EF which is turned on
    at time $t_{warm} = 10$~ps. The red line refers to the oscillatory field with period of
    10~ps. The green line is of 40~ps; and the blue of 200~ps.}
  \label{fig:tmp8}
\end{figure}
Fig.~\ref{fig:tmp8} (a)
presents the $x$-component of the molecular dipole moment as
a function of time. The red, green and blue lines report the
results of $\period =10$~ps, 40~ps and 200~ps, respectively. The black line
shows the dipole moment under constant EF
for reference.
It shows that the oscillating period of the molecular dipole moment
is the same as the period of the oscillatory EF. Therefore,
the molecular dipole moment
is able to respond to the external EF almost immediately.
At $\period =200$~ps, the maximum magnitude of the molecular
dipole moment is close to the value obtained in the case of constant EF, which
means that the variation of the oscillatory EF is so slow that the
alanine has enough time to 
relax its dipole. However,
for $\period =10$ and 40~ps, the EF oscillates faster and the molecule 
does not have enough time to fully relax the dipole, as a consequence the maximum dipole moment is smaller than the previous case.
Fig.~\ref{fig:tmp8}~(b) presents the orientational order
parameter. The notations are the same as Fig~\ref{fig:tmp8}~(a).
For all cases the order parameter is much
smaller than in the case of constant EF. One possible reason is that the
relaxation of the order parameter is very slow, and the molecule in the case of oscillatory field is
not exposed to a strong enough EF for a time long enough.
For  $\period =10$ and 40~ps, the orientation of the alanine shows a {very} weak tendency
to be parallel to the EF. It is worth to notice that for constant EF,
we also observe a quick alignment of the molecular orientation vector to the EF
on the time scale of 10~ps. For $\period =200$~ps, the molecule tends periodically
to be more perpendicular to the EF, but this directional preference is much
weaker than in the case of constant EF.

\begin{figure*}
  \centering
  \includegraphics[width=0.32\textwidth]{fig-meta-mode2-0010.eps}
  \includegraphics[width=0.32\textwidth]{fig-meta-mode2-0040.eps}
  \includegraphics[width=0.32\textwidth]{fig-meta-mode2-0200.eps}
  % \includegraphics[width=0.49\textwidth]{fig-meta-mode2-2000.eps}
  \caption{The probability of being in a given of conformation for alanine in a   periodically oscillatory EF. Different periods,
    i.e. $\period =10$~ps, 40~ps and 200~ps, are considered here.  The
    probability of being in a certain conformation $I$, which is
    denoted by $P_I$ in this paper, is plotted by colored lines against
    time. Red line: $I = \confaa$,  green line: $I = \confab$,  dark
    blue line: $I = \confba$,  pink line: $I = \confbb$ and  light blue
    line: $I = \confc$. For $\period =10$~ps and 40~ps, the nonequilibrium
    simulations last 1000~ps, while for $\period =200$~ps,
    they last 6400~ps.  }
  \label{fig:tmp9}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-1.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-2.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-3.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-4.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0010-iflux-5.eps}\\
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-1.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-2.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-3.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-4.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0040-iflux-5.eps}\\
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-1.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-2.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-3.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-4.eps}
  \includegraphics[width=0.19\textwidth]{fig-trans-mode2-0200-iflux-5.eps}\\
  \caption{
    Integrated probability flux $Q_{J,I}(t)$ for the periodically
    oscillatory EF, is plotted against time (horizontal axis, in picoseconds).
    The integrated
    probability flux is defined by $Q_{J,I}(t) = \int_0^t F_{J,I}(\tau) d\tau$, where
    $F_{J,I}(t)$ is the net probability flux from conformation $J$ to $I$.
    See also Eqn.~\eqref{eqn:tmp7} and~\eqref{eqn:tmp7a} for definitions
    of $F_{J,I}(t)$ and $Q_{J,I}(t)$, respectively.
    From top to bottom, rows report results with  period 10, 40 and 200~ps, respectively.
    From left to right, the five
    columns show the integrated flux $Q_{J,\confaa}$, $Q_{J,\confab}$,
    $Q_{J,\confba}$, $Q_{J,\confbb}$ and $Q_{J,\confc}$, respectively. In each plot,
    the red line stands for $J=\confaa$, green for $J=\confab$, dark blue for $J=\confba$,
    purple for $J=\confbb$ and light blue for $J=\confc$.
    An increasing value of $Q_{J,I}$ indicates a net probability
      flux from conformation $J$ to $I$,
      while a decreasing value of $Q_{J,I}$ indicates a net probability
      flux from $I$ to $J$.
  }
  \label{fig:tmp10}
\end{figure*}
